\chapter{Toolbox Functions}

\section{Calling Convention}

All Toolbox functions are long call functions using the Calypsi ``simple call'' calling convention:
\begin{itemize}
	\item left-most parameter is placed in D0.
	\item the remaining parameters are pushed on to the stack in right to left order (that is, the second parameter in a call is at the top of the stack just before the \verb+JSL+).
	\item 8-bit types are pushed as 16-bit values to avoid switching register sizes mid-call
	\item 24-bit types are pushed as 32-bit values for the same reason
	\item the return value is placed in the accumulator for 8 and 16-bit types, or in the X register and accumulator for 24 and 32 bit types (most significant bits in the X register).
	\item The caller is responsible for removing the parameters from the stack (if any) after the call returns.
\end{itemize}

The Toolbox functions are accessed through a jump table located in the FA2560K2's RAM, starting at 0x00000400, right after the CPU vector addresses. Each entry is six bytes long, and the address of each function is called out in their detailed descriptions below.

NOTE: Calypsi's ``simple call'' convention is not the fastest way to pass parameters to functions, and it is not Calypsi's only calling convention.
There is also a calling convention that uses pseudo-registers in the direct page to pass parameters.
Unfortunately, the rules for which parameter goes where in direct page are rather involved.
While that convention is preferable when Calypsi is the only compiler involved, the Toolbox needs to allow for other development tools to be used.
The stack based convention is more likely to be supported by other compilers. So speed was traded for broader compatibility.

\input{general.tex}   	% Updated to 68000 Mark II
\input{channel.tex}   	% Updated to 68000 Mark II
\input{block.tex}     	% Updated to 68000 Mark II
\input{fsys.tex}      	% Updated to 68000 Mark II
\input{text.tex}		% Updated to 68000 Mark II
\input{interrupt.tex}	% Updated to 68000 Mark II
%% \input{iec.tex}